---
title: "HW3 - 2"
author: "Renjie Liu"
date: "2/14/2021"
output: html_document
---
1. Split sample into two random samples of sizes 70% and 30%.
```{r}
library(ggplot2)
library(mlbench)
library(dplyr)
data("BostonHousing")
set.seed(1234)

train_index <- sample(seq_len(nrow(BostonHousing)),size = floor(0.7*nrow(BostonHousing)))

train <- BostonHousing[train_index,]

test <- BostonHousing[-train_index,]
```

2.Perform principal components of numeric variables from the Boston Housing Data on training sample.
```{r}
train <- scale(train[,c(-4,-9)])

test <- scale(test[,c(-4,-9)])

pca_train <- princomp(train)
```


3.Generate Scree Plots and select number of components you would retain.
```{r}
VAF = cumsum(pca_train$sdev^2/sum(pca_train$sdev^2))

data.frame(n = 1:length(VAF),Variance.Account.For = VAF) %>% 
  ggplot(aes(x=n, y=Variance.Account.For)) +
  geom_line() +
  geom_point()+
  geom_text(aes(label=VAF),vjust = -1)

"I will choose 5 componets because the curve is becoming flat after 5 components. 5 components will have a VAF of around 85%"
```

4.Plot Component 1 loadings (x-axis) versus Component 2 loadings (y-axis). Use this plot to interpret and name the Components. Repeat this by plotting Component (1) separately versus all components you decided to retain from Step 3 (Component 3, Component 4 etc). Can you interpret each of the components you decide to retain. In case a component is not interpretable, note that.
```{r}
biplot(pca_train$scores[1:10,c(1,2)],pca_train$loadings[,c(1,2)])


"Component 1 has the variances of all the predictors. For Component 1 , nox, age, indus, and ptratio have large positive effect on component 1. dis,medv, and rm have large negative effect on component 1. The lower the component 1 the higher the house price. I think component 1 is a good measure of all the information for a house, especially for the house price. 

For Component 2, dis, ptratio have very large positive effect on component 1. Medv,age, and nox have very large negative effect on component 2. Component 2 also measures the house price, and also measures the distance of the house."
```

```{r}
biplot(pca_train$scores[1:10,c(1,3)],pca_train$loadings[,c(1,3)])

"component 3 has limited information for the variabes mentioned in component 1 and component 2. However, component 3 is heavily affected by crim, tax, zn, and b. Therefore, I believe that component 3 measures the crime rate for the house."
```

```{r}
biplot(pca_train$scores[1:10,c(1,4)],pca_train$loadings[,c(1,4)])
```
component 4 has limited information for variabels in component 1 ,2,3. Component 4 is largely impacted by ptratio. The ptratio measures the pupil-teacher ratio by town. Therefore, i think component 4 measures the educatioal quality for the house's town.


```{r}
biplot(pca_train$scores[1:10,c(1,5)],pca_train$loadings[,c(1,5)])
```
Component 5 has even less information than 3. It is largely affted by b, crim, zn, and tax. it is similar to component 3 because it is also heavily impacted by zn, crim, and tax,b. However, the most notable variable impact component 5 is b. Therefore , component 4 measure the blacks porpotition for the house town


5. Perform the following:
a.Show that Component loadings are orthogonal.
```{r}
round(t(pca_train$loadings)%*%pca_train$loadings,5)
"The matrix is a diagonal matrix that has non-zero values on the diagonal and zeroes off the diagonal. This matrix shows that the component loadings are orthogonal"
```

b.Show that Component scores are orthogonal.
```{r}
round(cov(pca_train$scores),4)
"The matrix is a diagonal matrix that has non-zero values on the diagonal and zeroes off the diagonal. This matrix shows that the component scores are orthogonal"
```

c. Perform Test validation of Principal Components solution.
i. For Test validation, you will have to
1. predict the component scores in the Test [ using the predict()
function in R and transform function in Python
```{r}
test_factor <- predict(pca_train,test)
head(test_factor)
```
2. matrix multiply the predicted component scores from (1) above
with transpose of component loadings you derived from training
data set from Step 2 above. Refer to Page 52 of Class Lecture for
Session 4 for details.
```{r}
multiplyed <- test_factor[,1:5] %*% t(pca_train$loadings[,1:5])
head(multiplyed)
```
d. Compute the Variance Account For (R2) in the Test sample. That yields a
measure of Test performance.
```{r}
cor(as.vector(test),as.vector(multiplyed))^2
#VAF of 0.82 for the test performance
```

e.[OPTIONAL] Rotate the component loadings using varimax rotation. Look at the Loadings from the varimax rotation. Does it yield any different Interpretation of the Principal Components?
```{r}
rotate <- varimax(pca_train$loadings)

biplot(pca_train$scores[1:10,c(1,2)],rotate$rotmat[,c(1,2)])
"Yes the component is very different from the abvoe biplot of component 1 and component 2"
```

