---
title: "HW5-2"
author: "Renjie Liu"
date: "3/7/2021"
output: html_document
---
#### 1. Use same train and test samples from logistic regression assignment (3); Dataset is : Diabetes.
#### 2. R user – recommend to use undummy dataset. Attached in this announcement.
```{r}
library(rpart)
library(MASS)
library(caret)
data <- read.csv('diabetes_preprocessed_nodummies.csv')
data <- data[,c(3,5:12,1,2,4,13:37)]
data[,10:37] <- data.frame(lapply(data[,10:37],as.factor))
data <- data[,c(37,1:36)]
```

#### 3. Run LDA using select predictors and predict class results for train and holdout:
```{r}
set.seed(1234)
train_index<- sample(nrow(data), size = 0.7*nrow(data))
train <- data[train_index,]
test <- data[-train_index,]

train <- as.data.frame(train)
test <- as.data.frame(test)


train <- train[,c(1,2:19,22:28,31:32,36:37)]
test <- test[,c(1,2:19,22:28,31:32,36:37)]

lda.model = lda(train$readmitted~.,data=train)
lda.train.pred <-predict(lda.model)$class 
lda.test.pred <- predict(lda.model, newdata = test)$class
```
#### 4. Run QDA and predict class results for train and holdout:
```{r}

qda.model <- qda(readmitted~.,data=train)
qda.train.pred <-predict(qda.model)$class
qda.test.pred <- predict(qda.model, newdata = test)$class
```
#### 5. How often in train and test do LDA and QDA make the same prediction?
```{r}
confusionMatrix(lda.train.pred,qda.train.pred)
confusionMatrix(lda.test.pred,qda.test.pred)
# For the train set, The LDA and QDA will make the same prediction 82.2%.
# For the test set, The LDA and QDA will make the same prediction 82.4%.

```
#### 6.Compare confusion matrix for Train and holdout for LDA results for validation. Do the same for QDA results. Ignore ‘ensemble’ from last question. It is a typo; not needed for this assignment.
```{r}
confusionMatrix(lda.train.pred,train[,1])
confusionMatrix(lda.test.pred,test[,1])
# For the lda train, the model has a accuracy of 63%
# For the lda test, the model has a accuracy of 62%

confusionMatrix(qda.train.pred,train[,1])
confusionMatrix(qda.test.pred,test[,1])
# For the qda train, the model has a accuracy of 62%
# For the qda test, the model has a accuracy of 61%

# Based on the accuracy, the lda is better to predict for this dataset, but still the accuracy is not very high. 
```









